R Session Info

### 2018-02-11 analysis/RsessionInfo.Rmd
Files commited by wflow_commit.

---
title: "the choice of softImpte lambda"
author: "Wei Wang"
date: 2018-01-27
output: 
  html_document:
    code_folding: hide
---

@@

## Document

Based on the document of the authors ![https://cran.r-project.org/web/packages/softImpute/softImpute.pdf], we should choose lambda ($\lambda$) as "Ideally lambda should be chosen so that the solution reached has rank slightly
less than rank.max."

## Brease Cancer data in flash paper

```{r}
CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD"){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = fold)
  rowindex = matrix(sample(N,N),ncol = fold)

  missing= array(0,dim = c(fold,N,P))
  foldindex = array(0,dim = c(fold,fold,2))
  for(i in 1:fold){
    for(j in 1:fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% fold
    }
  }
  foldindex[which(foldindex == 0)] = fold
  for(i in 1:fold){
    missing[i, , ] = Y
    for(j in 1:fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
  }
  # c_s is the candicate of shrinkage parameter
  n_s = length(c_s)
  # rmse for each grids
  CVRMSE = rep(0,n_s)
  minrmse = Inf
  opt_s = 0
  # for each candidate, we run it N_sim times
  for(t_s in 1:n_s){
    # for each grid
    # each time we set the rmse to zeros
    rmse = rep(0,fold)
    for(i in 1:fold){
      if(method == "PMD"){
        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
      }else{
        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
      }
      if(length(out$d)==1){
        misshat = (out$d) * out$u %*% t(out$v)
      }else{
        misshat = out$u %*%  diag(out$d) %*% t(out$v)
      }
      for(j in 1:fold){
        # for each fold j
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
      }
    } #get the result for one run
    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }
  return(list(opt_s = opt_s, output = CVRMSE))
}

softImpute.wrapper = function(Y,ngrids = 10,K,fold = 10){
  library(softImpute)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0,100,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
  return(list(d = out$d, u = out$u, v = out$v))
}

OCV_index=function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold

  return(list(foldindex = foldindex, rowindex = rowindex, colindex = colindex))
}
# OCVindex = OCV_index(Y,k_fold = 5)

OCV_data = function(Y,OCVindex,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing= array(0,dim = c(k_fold,N,P))
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
  }
  return(missing)
}
# OCVdata = OCV_data(Y,OCVindex,k_fold = 5)

OCV_SSE = function(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50){
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing = OCVdata
  SSE = rep(0,k_fold)
  for(i in 1:k_fold){
    miss_hat = call_method(missing[i,,], method = method, Kmax = Kmax)
    for(j in 1:k_fold){
      SSE[i] = SSE[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
    }
  }
  RMSE = sqrt(sum(SSE)/(dim(Y)[1] * dim(Y)[2]))
  return(RMSE)
}

call_method = function(Y_data,method,Kmax = 50){
  if(method == "flash"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_wn"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_gb"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
    f = flashr2::flash_backfit(data,f_greedy)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "flash_bf"){
    data = flashr2::flash_set_data(Y_data)
    f_data = flashr2::flash_add_factors_from_data(data,K = Kmax)
    f = flashr2::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "PMD"){
    res_log = capture.output({out = PMA.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10)})
    if(length(out$d)==1){
      Y_hat =  (out$d) * out$u %*% t(out$v)
    }else{
      Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
    }
  }else if(method == 'PN'){
    library(ebnm)
    library(flashr2)
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == 'PMD_d'){
    out = PMA::PMD(Y_data, K = Kmax)
    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
  }else if(method == 'softImpute_d'){
    gsoft = softImpute::softImpute(Y_data, rank.max = Kmax)
    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  }else if(method == "softImpute"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    if(length(gsoft$d)==1){
      Y_hat =  try((gsoft$d) * gsoft$u %*% t(gsoft$v))
    }else{
      Y_hat =  try(gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v))
    }
  }else if(method == "SF_flash"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    LL = gsoft$u %*% diag(gsoft$d)
    FF = gsoft$v
    data = flashr2::flash_set_data(Y_data)
    f_data = flashr2::flash_add_lf(data, LL=LL, FF=FF)
    f = flashr2::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else{
    # stop("the method is not implemented yet, please check it out")
    Y_hat = matrix(0,ncol = dim(Y_data)[2],nrow = dim(Y_data)[1])
  }
  return(Y_hat)
}

```

read the data

```{r}
library(R.matlab)
## run the code
Y_centered = readMat("../data/output/missingdata/BreastCancer/example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
```

first we try to get the lambda based on the authors suggestion.

```{r}
OCVindex = OCV_index(Y,k_fold = 10)
OCVdata = OCV_data(Y,OCVindex,k_fold = 10)

colindex = OCVindex$colindex
rowindex = OCVindex$rowindex
foldindex = OCVindex$foldindex
missing = OCVdata
Y_data = missing[1,,]

out = softImpute(Y_data, rank.max = 40,lambda = 80)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 30)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 20)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 10)
length(out$d)
out = softImpute(Y_data, rank.max = 40,lambda = 9)
length(out$d)
```

so it seems 10 or 11 should be a good choice which "Ideally lambda should be chosen so that the solution reached has rank slightly less than rank.max"

## OCV choice

for the grids we choose in our paper, we use 0-100 with ten grids

```{r}
library(softImpute)
N = dim(Y_data)[1]
P = dim(Y_data)[2]
ngrids = 10
fold = 10
K = 40
c_s = seq(0,100,len=ngrids)
cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
cvout
```

the result fit the authors' suggestion well (we choose $\lambda = 11$).

But can we try smaller grids with OCV

```{r}
library(softImpute)
N = dim(Y_data)[1]
P = dim(Y_data)[2]
ngrids = 10
fold = 10
K = 40
c_s = seq(0,15,len=ngrids)
cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
cvout
```

At this time, we can see that we choose $\lambda = 5$ this time.

## check the result

```{r}
out = softImpute(Y_data, rank.max = 40,lambda = 10)
miss_hat = out$u %*% diag(out$d) %*% t(out$v)
SSE = 0
for(j in 1:fold){
      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
}
sqrt(SSE/sum(is.na(Y_data)))
```

```{r}
out = softImpute(Y_data, rank.max = 40,lambda = 5)
miss_hat = out$u %*% diag(out$d) %*% t(out$v)
SSE = 0
for(j in 1:fold){
      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
}
sqrt(SSE/sum(is.na(Y_data)))
```

we can see that the result is slightly improved.

>>>>

---
title: "R Session Info"
author: "Wei Wang"
date: 
output: html_document
---

@@

```{r}
library(flashr)
library(ssvd)
library(methods)
library(PMA)
library(R.matlab)
library(Matrix)
library(softImpute)
library(ggplot2)
```
